---
title: "Flegal Diff Misclassification From Non-diff Measurement"
author: "Chrystelle Kiang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```


The first goal is to re-create Flegal et al's simulation in "Differential Misclassification Arising from Nondifferential Errors in Exposure Measurement"  

E true exposure values, 100 samples evenly spaced over specified interval  
E' measured exposure values: true + error term, where Error terms ~N with mean 0, "specified" SD  
true exposure values range integers 1,600 - 2,499, high/low cutoff of 2,200  
standard deviation of measurement error varied at 100, 300, 500  

### Recreating simulations
Created a function to replicate the scenario from Flegal et al that outputs the Se and Sp overall and by disease status.
The first time I made this function, disease was being assigned anew each simulation except for exposure. (Is in a chunk later in code). To reduce variability, created function that will fix the exposure AND disease status before measurement error is simulated. Note though that Flegal et al. did not fix disease before simulations, since they present averages of the "true" and "observed" RRs (Table 3).  

Function arguments:
* N: sample size, uniformly distributed across range of exposure values
* distrib: "UNIF" or "NORM"
* Eparam1, Eparam2: if uniform, this is min/ max of exposure values. if normal, this is mean and SD.
* cutoff: the cutpoint for categorizing high/ low <cutoff
* a, b: parameters for the linear logistic probability of disease
* MEsigma: vector of the standard deviations of measurement error  
* Nsims: number of simulations 

```{r functionsV2} 
measured_flegal <- function(N, E, D, cutoff, MEsigma, Nsims){
  values <- matrix(nrow = Nsims, ncol = 8, dimnames = list(c(), c()))
  results <- matrix(nrow = length(MEsigma), ncol = 9, dimnames = list(c(),c("ME", "SeAll", "SeD1", "SeD0", "SpAll", "SpD1", "SpD0", "RRtrue", "RRobs")))
  
  Ehigh <- ifelse(E > cutoff, 1, 0)
  Eerror <- c()
  Eprime <- c()
  Eprimehigh <- c()
  
  for (j in seq_along(MEsigma)){
    
    for (i in 1:Nsims){  
      Eerror <- rnorm(N, mean = 0, sd = MEsigma[[i]])
      Eprime <- E + Eerror
      Eprimehigh <- ifelse(Eprime > cutoff, 1, 0)
      
      values[i,1] <- sum(Eprimehigh==1 & Ehigh==1)/sum(Ehigh==1) #Se all
      values[i,2] <- sum(Eprimehigh[D==1]==1 & Ehigh[D==1]==1)/sum(Ehigh[D==1]==1) #Se D=1
      values[i,3] <- sum(Eprimehigh[D==0]==1 & Ehigh[D==0]==1)/sum(Ehigh[D==0]==1) #Se D=0
      values[i,4] <- sum(Eprimehigh==0 & Ehigh==0)/sum(Ehigh==0) #Sp all
      values[i,5] <- sum(Eprimehigh[D==1]==0 & Ehigh[D==1]==0)/sum(Ehigh[D==1]==0) #Sp D=1
      values[i,6] <- sum(Eprimehigh[D==0]==0 & Ehigh[D==0]==0)/sum(Ehigh[D==0]==0) #Sp D=0
      values[i,7] <- (sum(D[Ehigh==1]==1)/sum(Ehigh==1))/(sum(D[Ehigh==0]==1)/sum(Ehigh==0)) #RR
      values[i,8] <- (sum(D[Eprimehigh==1]==1)/sum(Eprimehigh==1))/(sum(D[Eprimehigh==0]==1)/sum(Eprimehigh==0)) #RR measured
    }
    results[j,] <- cbind(MEsigma[[j]], mean(values[,1]), mean(values[,2]), mean(values[,3]), mean(values[,4]), mean(values[,5]), mean(values[,6]), mean(values[,7]), mean(values[,8]))
  }
  return(results)
}

# This function calls previous function. E is fixed, D is not.
# if distribution is UNIF, Eparam1 = E min, Eparam2 = E max
# if distribution is NORM, Eparam1 = E mean, Eparam2 = standard deviation of exposure
simulate_flegal_V2 <- function(N, distrib, Eparam1, Eparam2, cutoff, a, b, MEsigma, Nsims){
    output <- matrix(nrow = length(MEsigma), ncol = 17, 
                   dimnames = list(c(),c("N", "Distrib", "Eparam1", "Eparam2", "cutoff", "a", "b", "Nsims", "ME", "SeAll", "SeD1", "SeD0", "SpAll", "SpD1", "SpD0", "RRtrue","RRobs")))
  E <- c()
  D <- c()
  cutoff <- cutoff
  MEsigma <- MEsigma
  # initializing for each iteration of disease 
  one_d <- list()
  all_means <- matrix(nrow = length(MEsigma), ncol=9)

  for (i in 1:Nsims){
      if (distrib == "UNIF"){E <- as.integer(runif(N, min = Eparam1, max = Eparam2))}  
      if (distrib == "NORM"){E <- as.integer(rnorm(N, mean = Eparam1, sd = Eparam2))}
      if (distrib == "EVEN"){E <- as.integer(seq(from = Eparam1, to = Eparam2, length.out = N))}
    D <- rbinom(n = N, size = 1, p = plogis(a + b*E)) # this will be created Nsims times
    one_d[[i]] <- measured_flegal(N, E, D, cutoff, MEsigma, Nsims)
  }
  all_means <- Reduce("+", one_d)/ length(one_d)
  output <- as.data.frame(cbind(N, distrib, Eparam1, Eparam2, cutoff, a, b, Nsims, all_means))
  return(output)
}

newvar <- c(100, 300, 500)

simulate_flegal_V2(100, "EVEN", 2050, 150, 2200, -4.5, 0.0019, newvar, 200)

testcase2 <- simulate_flegal(1000, "NORM", 2050, 150, 2200, -4.5, 0.0019, newvar, 400)

```

```{r createfunctions} 
# if distribution is UNIF, Eparam1 = E min, Eparam2 = E max
# if distribution is NORM, Eparam1 = E mean, Eparam2 = standard deviation of exposure
simulate_flegal <- function(N, distrib, Eparam1, Eparam2, cutoff, a, b, MEsigma, Nsims){
  E <- c()
  D <- c()
  Ehigh <- c()
  Eerror <- c()
  Eprime <- c()
  Eprimehigh <- c()
  values <- matrix(nrow = length(MEsigma), ncol = 11, dimnames = list(c(),c("MEsd", "SeAll", "SeD1", "SeD0", "SpAll", "SpD1", "SpD0", "RRtrue", "RRobs", "E prev", "Eprime prev")))
  results <- list()
  all_means <- matrix(nrow = length(MEsigma), ncol=11)
  output <- matrix(nrow = length(MEsigma), ncol = 19, dimnames = list(c(),c("N", "Distrib", "Eparam1", "Eparam2", "cutoff", "a", "b", "Nsims", "ME", "SeAll", "SeD1", "SeD0", "SpAll", "SpD1", "SpD0", "RRtrue","RRobs", "Eprev", "Eprimeprev")))
  
  for (j in 1:Nsims){
    if (distrib == "UNIF"){E <- as.integer(runif(N, min = Eparam1, max = Eparam2))}  
    if (distrib == "NORM"){E <- as.integer(rnorm(N, mean = Eparam1, sd = Eparam2))}
    if (distrib == "EVEN"){E <- as.integer(seq(from = Eparam1, to = Eparam2, length.out = N))}
    D <- rbinom(n = N, size = 1, p = plogis(a + b*E)) 
    Ehigh <- ifelse(E > cutoff, 1, 0)
    
    for (i in seq_along(MEsigma)){
      Eerror <- rnorm(N, mean = 0, sd = MEsigma[i])
      Eprime <- E + Eerror
      Eprimehigh <- ifelse(Eprime > cutoff, 1, 0)
      
      values[i,1] <- MEsigma[i] #this will be constant but want it to pass through
      values[i,2] <- sum(Eprimehigh[Ehigh==1]==1)/sum(Ehigh==1) #Se all
      values[i,3] <- sum(Eprimehigh[D==1]==1 & Ehigh[D==1]==1)/sum(Ehigh[D==1]==1) #Se D=1
      values[i,4] <- sum(Eprimehigh[D==0]==1 & Ehigh[D==0]==1)/sum(Ehigh[D==0]==1) #Se D=0
      values[i,5] <- sum(Eprimehigh[Ehigh==0]==0)/sum(Ehigh==0) #Sp all
      values[i,6] <- sum(Eprimehigh[D==1]==0 & Ehigh[D==1]==0)/sum(Ehigh[D==1]==0) #Sp D=1
      values[i,7] <- sum(Eprimehigh[D==0]==0 & Ehigh[D==0]==0)/sum(Ehigh[D==0]==0) #Sp D=0
      values[i,8] <- (sum(D[Ehigh==1]==1)/sum(Ehigh==1))/(sum(D[Ehigh==0]==1)/sum(Ehigh==0)) #RR
      values[i,9] <- (sum(D[Eprimehigh==1]==1)/sum(Eprimehigh==1))/(sum(D[Eprimehigh==0]==1)/sum(Eprimehigh==0)) #RR measured
      values[i,10] <- sum(Ehigh == 1)/N
      values[i,11] <- sum(Eprimehigh == 1)/N #measured exposure prevalence
    }
    results[[j]] <- values
  }
  all_means <- Reduce("+", results)/ length(results)
  output <- as.data.frame(cbind(N, distrib, Eparam1, Eparam2, cutoff, a, b, Nsims, all_means))
  return(output)
}

```

*Case 1* Repeating the original simulation

So for the main simulations, they originally varied:  

* standard deviation of measurement error: 100, 300, 500  
* three sets of (a, b) parameters for disease: (-4.5, 0.0019), (-8.0, 0.0035), (-16.0, 0.0072)  
and did 200 simulations of each on the nine combinations.  

```{r repeatflegal}
# function arguments: N, distrib, Eparam1, Eparam2, cutoff, a, b, MEsigma, Nsims
# if distribution is UNIF or EVEN, Eparam1 = E min, Eparam2 = E max
# if distribution is NORM, Eparam1 = E mean, Eparam2 = standard deviation of exposure
# A, B, C reflect the different regression models/ RR
MEsigma <- c(100, 300, 500)
set.seed(303)
case1A <- simulate_flegal(100, "EVEN", 1600, 2499, 2200, -4.5, 0.0019, MEsigma, 200)
case1B <- simulate_flegal(100, "EVEN", 1600, 2499, 2200, -8.0, 0.0035, MEsigma, 200)
case1C <- simulate_flegal(100, "EVEN", 1600, 2499, 2200, -16.0, 0.0072, MEsigma, 200)

case1 <- rbind(case1A, case1B, case1C)
```

*Case 2* Repeating, but uniform distribution rather than even spacing
```{r, case2}
set.seed(303)
case2A <- simulate_flegal(100, "UNIF", 1600, 2499, 2200, -4.5, 0.0019, MEsigma, 200)
case2B <- simulate_flegal(100, "UNIF", 1600, 2499, 2200, -8.0, 0.0035, MEsigma, 200)
case2C <- simulate_flegal(100, "UNIF",1600, 2499, 2200, -16.0, 0.0072, MEsigma, 200)

case2 <- rbind(case2A, case2B, case2C)
```


*Case 3*  
Increasing the number of simulations from 200 to 1,000  
```{r case3}
# function arguments: N, distrib, Eparam1, Eparam2, cutoff, a, b, MEsigma, Nsims
set.seed(303)
case3A <- simulate_flegal(100, "UNIF", 1600, 2499, 2200, -4.5, 0.0019, MEsigma, 1000)
case3B <- simulate_flegal(100, "UNIF", 1600, 2499, 2200, -8.0, 0.0035, MEsigma, 1000)
case3C <- simulate_flegal(100, "UNIF", 1600, 2499, 2200, -16.0, 0.0072, MEsigma, 1000)

case3 <- rbind(case3A, case3B, case3C)
```

*Case 4*  
Increasing number of samples 
```{r case4, eval = F}
set.seed(303)
case4A <- simulate_flegal(10000, "UNIF", 1600, 2499, 2200, -4.5, 0.0019, MEsigma, 1000)
case4B <- simulate_flegal(10000, "UNIF", 1600, 2499, 2200, -8.0, 0.0035, MEsigma, 1000)
case4C <- simulate_flegal(10000, "UNIF", 1600, 2499, 2200, -16.0, 0.0072, MEsigma, 1000)

case4 <- rbind(case4A, case4B, case4C)
```



Range under uniform distribution is 1,600 to 2,499. 
If mean = midpoint of 2050, SD of 150 would make most of the values fall within 1,600 - 2,5000 range.    

Normally distributed exposure:  
*Case 5* Exposure distribution centered about cutpoint   
*case 6* Changing the exposure distribution to centered about midpoint ~N(2050, 150)   
*Case 7* Rare exposure (mean is lower than midpoint)   

```{r normals}
# function arguments: N, distrib, Eparam1, Eparam2, cutoff, a, b, MEsigma, Nsim
# if distribution is UNIF, Eparam1 = E min, Eparam2 = E max
# if distribution is NORM, Eparam1 = E mean, Eparam2 = standard deviation of exposure

MEsigma <- c(100, 300, 500)

set.seed(303)
case5A <- simulate_flegal(1000, "NORM", 2200, 150, 2200, -4.5, 0.0019, MEsigma, 1000)
case5B <- simulate_flegal(1000, "NORM", 2200, 150, 2200, -8.0, 0.0035, MEsigma, 1000)
case5C <- simulate_flegal(1000, "NORM", 2200, 150, 2200, -16.0, 0.0072, MEsigma, 1000)
case5 <- rbind(case5A, case5B, case5C)

##### Other ~N
# Case 6: Normal distribution around the cutpoint (mean 2,200)  
# Case 7: exposure ~N but 'high' is rare (mean 1,900)
# Case 8: big sample size
set.seed(303)
case6A <- simulate_flegal(1000, "NORM", 2050, 150, 2200, -4.5, 0.0019, MEsigma, 1000)
case6B <- simulate_flegal(1000, "NORM", 2050, 150, 2200, -8.0, 0.0035, MEsigma, 1000)
case6C <- simulate_flegal(1000, "NORM", 2050, 150, 2200, -16.0, 0.0072, MEsigma, 1000)
case6 <- rbind(case6A, case6B, case6C)

set.seed(303)
case7A <- simulate_flegal(1000, "NORM", 1900, 150, 2200, -4.5, 0.0019, MEsigma, 1000)
case7B <- simulate_flegal(1000, "NORM", 1900, 150, 2200, -8.0, 0.0035, MEsigma, 1000)
case7C <- simulate_flegal(1000, "NORM", 1900, 150, 2200, -16.0, 0.0072, MEsigma, 1000)
case7 <- rbind(case7A, case7B, case7C)
```


```{r more, eval = FALSE}
set.seed(303)
case8A <- simulate_flegal(10000, "NORM", 2050, 150, 2200, -4.5, 0.0019, MEsigma, 1000)
case8B <- simulate_flegal(10000, "NORM", 2050, 150, 2200, -8.0, 0.0035, MEsigma, 1000)
case8C <- simulate_flegal(10000, "NORM", 2050, 150, 2200, -16.0, 0.0072, MEsigma, 1000)

case8 <- rbind(case8A, case8B, case8C)
```


```{r savestuff, include = FALSE, eval = FALSE}
# at this point, I typically export as csv but with uniform cases in one grouping and normal separately, since the variables retained are slightly different. I then create an xlsx to combine
all_cases <- rbind(case1, case2, case3, case4, case5, case6, case7)
write.csv(all_cases, "flegal_export_132022.csv")
```


#### Recreating original figures  

E true exposure values, 100 samples evenly spaced over specified interval  
E' measured exposure values: true + error term, where Error terms ~N with mean 0, "specified" SD  
true exposure values range integers 1,600 - 2,499, high/low cutoff of 2,200  
standard deviation of measurement error varied at 100, 300, 500  

They first look at misclassification close to cutpoints of 2,000 and 2,200. Figure 1 plots the measured exposure values (Y axis) against the true exposure values (X axis) at two cutpoints.

```{r fig1}
pop0 <- tibble(E = as.integer(seq(from = 1600, to = 2499, length.out = 100)), 
               Eerror = rnorm(100, mean = 0, sd = 150), 
               Eprime = E + Eerror, 
               E_high = ifelse(E > 2200, 1, 0), 
               Eprime_high = ifelse(Eprime > 2200,1, 0),
               E_mis = ifelse(E_high != Eprime_high, 1, 0), 
               D = rbinom(n = 100, size = 1, p = plogis(-10 + 0.004*E))
)

sp1 <- ggplot(pop0, aes(x = E, y = Eprime)) + geom_point() + scale_x_continuous(limits = c(1400, 2600), breaks = seq(1400,2600, 200)) + scale_y_continuous(limits = c(1200, 2800), breaks = seq(1200,2800, 200)) + theme_bw()

Figure1a <- sp1 + labs(title = "Cutpoint of 2,000", x = "True Value", y = "Measured Value") + geom_vline(xintercept = 2000) + geom_hline(yintercept = 2000, linetype = "dashed") 
# Figure1a

Figure1b <- sp1 + labs(title = "Original scenario (n = 100)", x = "True Value", y = "Measured Value") + geom_vline(xintercept = 2200) + geom_hline(yintercept = 2200, linetype = "dashed")
# Figure1b
```

### Probability of misclassification and probability of disease
p probability of disease for each true exposure value, linear logistic model  ln(p/1-p) = a + b*E  
parameters a, b  
disease status 0 or 1 based on binomial n=1, prob of success = p from above  
Probability of misclassification is based on cumulative normal distribution. I based it on the difference between true value of E and the cutpoint; expected difference is 0.  

Figure 2 shows:  

1. expected probability of disease for a linear logistic model (a = -10, b = 0.004) AND  
2. probability of misclassification into low- and high- exposure based on normally distributed exposure measurement error (mean = 0, standard deviation = 150) 

Figure 3 shows: 

1. the expected probability of disease p as a quadratic function of the true exposure value E (p = 6.592 - 0.00673.E + 0.00000112E^2) over the specified range of values AND  
2. probability of misclassification (as in fig 2) 

```{r probD}
# expected probability of disease 
prob_D <- glm(D ~ E, data = pop0, family = binomial)
expected_pD <- predict(prob_D, pop0, type = "response")

# probability of misclassification
pop0$E_diff <- abs(pop0$E-2200)
pop0$a <- 1 - pnorm(pop0$E_diff, 0, 150)
# ggplot(pop, aes(x = E, y = a))+ geom_line() + geom_vline(xintercept = 2200) + theme_bw()

# Figure 2
flegal_figure2 <- ggplot(pop0, aes(x = E)) + geom_line(aes(y = expected_pD)) + geom_line(aes(y = a), linetype = 2) + geom_vline(xintercept = 2200) + theme_bw() + labs(title = "Fig 2. Prob of misclassification (dashed) & prob of disease (solid)", x = "True Value", y = "Probability")
flegal_figure2

# expected probability of disease p as a quadratic function of E
pop0$expected_pD_quad <- 6.592 - 0.00673*pop0$E + 0.00000172*(pop0$E**2)

# Figure 3:
flegal_figure3 <- ggplot(pop0, aes(x = E)) + geom_line(aes(y = expected_pD_quad)) + geom_line(aes(y = a), linetype = 2) + geom_vline(xintercept = 2200) + theme_bw() + labs(title = "Fig 3. Prob of misclassification (dashed) & prob of disease, quadratic term (solid)", x = "True Value", y = "Probability")

# probability of misclassification, varied SD to match simulation conditions
pop0$a100 <- 1 - pnorm(pop0$E_diff, 0, 100)
pop0$a300 <- 1 - pnorm(pop0$E_diff, 0, 300)
pop0$a500 <- 1 - pnorm(pop0$E_diff, 0, 500)

ggplot(pop0, aes(x = E)) + geom_line(aes(y = expected_pD), color = "darkgrey") + geom_line(aes(y = a), linetype = 2) + geom_line(aes(y = a100), color = "green") + geom_line(aes(y = a300), color = "green3") + geom_line(aes(y = a500), color = "darkgreen") + geom_vline(xintercept = 2200) + theme_bw() + labs(title = "Prob of misclassification & prob of disease (solid)", x = "True Value", y = "Probability")

```
The green lines are the expected probabilities of misclassification. 

Flegal et al. plot the probability of misclassification as a function of the distribution of the measurement error; not sure that it is actually the probability of misclassification? 
Why did they do that?? I think it may be more appropriate to simulate it and maybe re-create the population 100 times then get the proportion who were classified for each value of the exposure. Idk.

```{r morefunctions, eval = F}
uni_sim <- function(N, Emin, Emax, cutpt, a, b, MEsigma, Nsims){
  
  E <- as.integer(runif(N, min = Emin, max = Emax))  
  values <- matrix(nrow = Nsims, ncol = 8, dimnames = list(c(), c("Se", "SeD1", "SeD0", "Sp", "SpD1", "SpD0", "RRtrue", "RRobs")))
  results <- matrix(nrow = length(MEsigma), ncol = 16, dimnames = list(c(),c("N", "Emin", "Emax", "cutoff", "a", "b", "Nsims", "ME", "SeAll", "SeD1", "SeD0", "SpAll", "SpD1", "SpD0", "RRtrue", "RRobs")))
  
  
  for (j in seq_along(MEsigma)){
    
    for (i in 1:Nsims){
      D <- rbinom(n = N, size = 1, p = plogis(a + b*E)) 
      Eerror <- rnorm(N, mean = 0, sd = MEsigma)
      Eprime <- E + Eerror
      Ehigh <- ifelse(E > cutpt, 1, 0)
      Eprimehigh <- ifelse(Eprime > cutpt, 1, 0)
      
      values[i,1] <- sum(Eprimehigh==1 & Ehigh==1)/sum(Ehigh==1)
      values[i,2] <- sum(Eprimehigh[D==1]==1 & Ehigh[D==1]==1)/sum(Ehigh[D==1]==1)
      values[i,3] <- sum(Eprimehigh[D==0]==1 & Ehigh[D==0]==1)/sum(Ehigh[D==0]==1)
      
      values[i,4] <- sum(Eprimehigh==0 & Ehigh==0)/sum(Ehigh==0)
      values[i,5] <- sum(Eprimehigh[D==1]==0 & Ehigh[D==1]==0)/sum(Ehigh[D==1]==0)
      values[i,6] <- sum(Eprimehigh[D==0]==0 & Ehigh[D==0]==0)/sum(Ehigh[D==0]==0)
      
      values[i,7] <- (sum(D==1 & Ehigh==1)/sum(Ehigh==1))/(sum(D==1 & Ehigh==0)/sum(Ehigh==0))
      values[i,8] <- (sum(D==1 & Eprimehigh==1)/sum(Eprimehigh==1))/(sum(D==1 & Eprimehigh==0)/sum(Eprimehigh==0))
      
    }
    #  "N", "Emin", "Emax", "cutoff", "a", "b", "Nsims", "ME", "Se", "SeD1", "SeD0", "Sp", "SpD1", "SpD0", "RRtrue", "RRobs"
    results[j,] <- cbind(N, Emin, Emax, cutpt, a, b, Nsims, MEsigma[[j]], mean(values[,1]), mean(values[,2]), mean(values[,3]), mean(values[,4]), mean(values[,5]), mean(values[,6]), mean(values[,7]), mean(values[,8]))
  }
  
  results <- as.data.frame(results)
  return(results)
  
}

# Normal
normal_sim <- function(N, Emean, Esigma, cutpt, a, b, MEsigma, Nsims){
  E <- rnorm(N, mean = Emean, sd = Esigma)
  
  values <- matrix(nrow = Nsims, ncol = 8, dimnames = list(c(), c("Se", "SeD1", "SeD0", "Sp", "SpD1", "SpD0", "RRtrue", "RRobs")))
  
  params <- matrix(nrow = length(MEsigma), ncol = 16, dimnames = list(c(),c("N", "Emean", "ESD", "cutoff", "a", "b", "Nsims", "ME", "Se", "SeD1", "SeD0", "Sp", "SpD1", "SpD0", "RRtrue", "RRobs")))
  
  for (j in seq_along(MEsigma)){
    
    for (i in 1:Nsims){
      D <- rbinom(n = N, size = 1, p = plogis(a + b*E))
      Eerror <- rnorm(N, mean = 0, sd = MEsigma)
      Eprime <- E + Eerror
      Ehigh <- ifelse(E > cutpt, 1, 0)
      Eprimehigh <- ifelse(Eprime > cutpt, 1, 0)
      
      values[i,1] <- sum(Eprimehigh==1 & Ehigh==1)/sum(Ehigh==1)
      values[i,2] <- sum(Eprimehigh[D==1]==1 & Ehigh[D==1]==1)/sum(Ehigh[D==1]==1)
      values[i,3] <- sum(Eprimehigh[D==0]==1 & Ehigh[D==0]==1)/sum(Ehigh[D==0]==1)
      
      values[i,4] <- sum(Eprimehigh==0 & Ehigh==0)/sum(Ehigh==0)
      values[i,5] <- sum(Eprimehigh[D==1]==0 & Ehigh[D==1]==0)/sum(Ehigh[D==1]==0)
      values[i,6] <- sum(Eprimehigh[D==0]==0 & Ehigh[D==0]==0)/sum(Ehigh[D==0]==0)
      
      values[i,7] <- (sum(D==1 & Ehigh==1)/sum(Ehigh==1))/(sum(D==1 & Ehigh==0)/sum(Ehigh==0))
      values[i,8] <- (sum(D==1 & Eprimehigh==1)/sum(Eprimehigh==1))/(sum(D==1 & Eprimehigh==0)/sum(Eprimehigh==0))
    }
    
    params[j,] <- cbind(N, Emin, Emax, cutpt, a, b, Nsims, MEsigma[[j]], mean(values[,1]), mean(values[,2]), mean(values[,3]), mean(values[,4]), mean(values[,5]), mean(values[,6]), mean(values[,7]), mean(values[,8]))
    
  }
  
  params <- as.data.frame(params) 
  return(params)
}

flegal_sim <- function(N, Emin, Emax, cutpt, a, b, MEsigma, Nsims){
  
  E <- as.integer(seq(from = Emin, to = Emax, length.out = N))
  values <- matrix(nrow = Nsims, ncol = 8, dimnames = list(c(), c("Se", "SeD1", "SeD0", "Sp", "SpD1", "SpD0", "RRtrue", "RRobs")))
  results <- matrix(nrow = length(MEsigma), ncol = 16, dimnames = list(c(),c("N", "Emin", "Emax", "cutoff", "a", "b", "Nsims", "ME", "SeAll", "SeD1", "SeD0", "SpAll", "SpD1", "SpD0", "RRtrue", "RRobs")))
  
  allresults <- data.frame()
  
  
  for (j in seq_along(MEsigma)){
    
    for (i in 1:Nsims){
      D <- rbinom(n = N, size = 1, p = plogis(a + b*E)) 
      Eerror <- rnorm(N, mean = 0, sd = MEsigma)
      Eprime <- E + Eerror
      Ehigh <- ifelse(E > cutpt, 1, 0)
      Eprimehigh <- ifelse(Eprime > cutpt, 1, 0)
      
      values[i,1] <- sum(Eprimehigh==1 & Ehigh==1)/sum(Ehigh==1)
      values[i,2] <- sum(Eprimehigh[D==1]==1 & Ehigh[D==1]==1)/sum(Ehigh[D==1]==1)
      values[i,3] <- sum(Eprimehigh[D==0]==1 & Ehigh[D==0]==1)/sum(Ehigh[D==0]==1)
      
      values[i,4] <- sum(Eprimehigh==0 & Ehigh==0)/sum(Ehigh==0)
      values[i,5] <- sum(Eprimehigh[D==1]==0 & Ehigh[D==1]==0)/sum(Ehigh[D==1]==0)
      values[i,6] <- sum(Eprimehigh[D==0]==0 & Ehigh[D==0]==0)/sum(Ehigh[D==0]==0)
      
      values[i,7] <- (sum(D==1 & Ehigh==1)/sum(Ehigh==1))/(sum(D==1 & Ehigh==0)/sum(Ehigh==0))
      values[i,8] <- (sum(D==1 & Eprimehigh==1)/sum(Eprimehigh==1))/(sum(D==1 & Eprimehigh==0)/sum(Eprimehigh==0))
    }
    # "N", "Emin", "Emax", "cutoff", "a", "b", "Nsims", "ME", "SeAll", "SeD1", "SeD0", "SpAll", "SpD1", "SpD0", "RRtrue", "RRobs"
    results[j,] <- cbind(N, Emin, Emax, cutpt, a, b, Nsims, MEsigma[[j]], mean(values[,1]), mean(values[,2]), mean(values[,3]), mean(values[,4]), mean(values[,5]), mean(values[,6]), mean(values[,7]), mean(values[,8]))
    
  }
  
  allresults <- as.data.frame(results)
  return(allresults)
}
```


This is the first working one that does not have disease fixed across simulations- used to create results shared Aug 10 with group
```{r function1}
flegal_sim <- function(N, min, max, cutpt, a, b, sigma, sims){
  # True exposure
  E <- seq(from = min, to = max, length.out = N) 
  Se_all <- c()
  Se_D1 <- c()
  Se_D0 <- c()
  Sp_all <- c()
  Sp_D1 <- c()
  Sp_D0 <- c()
  RR_true <- c()
  RR_obs <- c()
  
  for (i in 1:sims){
    # set up
    Eerror <- rnorm(N, mean = 0, sd = sigma)
    Eprime <- E + Eerror
    E_high <- ifelse(E > cutpt, 1, 0)
    Eprime_high <- ifelse(Eprime > cutpt,1, 0)
    D <- rbinom(n = N, size = 1, p = plogis(a + b*E)) 
    
    # storing Se, Sp, RR values for each simulation
    Se_all[i] <- sum(Eprime_high==1 & E_high==1)/sum(E_high==1)
    Se_D1[i] <- sum(Eprime_high[D==1]==1 & E_high[D==1]==1)/sum(E_high[D==1]==1)
    Se_D0[i] <- sum(Eprime_high[D==0]==1 & E_high[D==0]==1)/sum(E_high[D==0]==1)
    
    Sp_all[i] <- sum(Eprime_high==0 & E_high==0)/sum(E_high==0)
    Sp_D1[i] <- sum(Eprime_high[D==1]==0 & E_high[D==1]==0)/sum(E_high[D==1]==0)
    Sp_D0[i] <- sum(Eprime_high[D==0]==0 & E_high[D==0]==0)/sum(E_high[D==0]==0)
    
    RR_true[i] <- ((sum(D==1 & E_high==1)/sum(E_high==1))/(sum(D==1 & E_high==0)/sum(E_high==0)))
    RR_obs[i] <- ((sum(D==1 & Eprime_high==1)/sum(Eprime_high==1))/(sum(D==1 & Eprime_high==0)/sum(Eprime_high==0)))
  }
  
  params <- data.frame(N, cutpt, a, b, sims, sigma, mean(Se_all), mean(Se_D1), mean(Se_D0), mean(Sp_all), mean(Sp_D1), mean(Sp_D0), mean(RR_true), mean(RR_obs))
  
  colnames(params) <- c("N", "cutoff", "a", "b", "Nsim", "SD", "Se all", "Se D=1", "Se D=0", "Sp all", "Sp D=1", "Sp D=0", "RR true", "RR obs")
  return(params)
}

set.seed(30308)

# case 1 is repeating the simulations from the paper 
# A. a = -4.5, b = 0.0019 
scenarioA1_100 <- flegal_sim(200, 1600, 2499, 2200, -4.5, 0.0019, 100, 200)
scenarioA1_300 <- flegal_sim(200, 1600, 2499, 2200, -4.5, 0.0019, 300, 200)
scenarioA1_500 <- flegal_sim(200, 1600, 2499, 2200, -4.5, 0.0019, 500, 200)
# B. a = -8.0, b = 0.0035
scenarioB1_100 <- flegal_sim(200, 1600, 2499, 2200, -8.0, 0.0035, 100, 200)
scenarioB1_300 <- flegal_sim(200, 1600, 2499, 2200, -8.0, 0.0035, 300, 200)
scenarioB1_500 <- flegal_sim(200, 1600, 2499, 2200, -8.0, 0.0035, 500, 200)
# C. a = -16.0, b = 0.0072
scenarioC1_100 <- flegal_sim(200, 1600, 2499, 2200, -16.0, 0.0072, 100, 200)
scenarioC1_300 <- flegal_sim(200, 1600, 2499, 2200, -16.0, 0.0072, 300, 200)
scenarioC1_500 <- flegal_sim(200, 1600, 2499, 2200, -16.0, 0.0072, 500, 200)
scenario1 <- rbind(scenarioA1_100, scenarioA1_300, scenarioA1_500, scenarioB1_100, scenarioB1_300, scenarioB1_500, scenarioC1_100, scenarioC1_300, scenarioC1_500)
```


